---
title: "Lab 7: Inference for Numerical Data"
author: "Julian Adames-Ng"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
library(infer)

data('yrbss', package='openintro')

yrbss

nrow(yrbss)

```

### Exercise 1

What are the cases in this data set? How many cases are there in our sample?

-Each row corresponds to a case. There are 13583 cases in this data set.

```{r code-Exercise2}
summary(yrbss$weight)
```

### Exercise 2

How many observations are we missing weights from?

-There are 1004 NA values which correspond to missing weights.

...

```{r code-Exercise3}

yrbss <- yrbss %>% 
  mutate(physical_3plus = ifelse(yrbss$physically_active_7d > 2, "yes", "no"))

yrbss

yrbss %>%
  group_by(physical_3plus) %>%
  summarise(mean_weight = mean(weight, na.rm = TRUE))

yrbss$physical_3plus <- as.factor(yrbss$physical_3plus)

yrb_plot <- yrbss %>%
  filter(!is.na(weight),!is.na(physical_3plus))
p <- ggplot(yrb_plot, aes(x = physical_3plus,y = weight))+
  geom_boxplot()
p

```

### Exercise 3

Make a side-by-side boxplot of physical_3plus and weight. Is there a relationship between these two variables? What did you expect and why?

-

...

```{r code-Exercise4}
summary(yrbss$weight)
```

### Exercise 4

How many observations are we missing weights from?

-There are 1004 NA values which correspond to missing weights.

...

```{r code-Exercise5}

yrbss <- yrbss %>% 
  mutate(physical_3plus = ifelse(yrbss$physically_active_7d > 2, "yes", "no"))

yrbss

```

### Exercise 5

Write the hypotheses for testing if the average weights are different for those who exercise at least times a week and those who don’t.

-Null: Students who are active 3 or more days per week have the SAME mean weight as students who are not active 3 or more days per week.

-Alternative: Students who are active 3 or more days per week have a different mean weight as students who are not active 3 or more days per week


```{r code-Exercise6}

obs_diff <- yrbss %>%
  specify(weight ~ physical_3plus) %>%
  calculate(stat = "diff in means", order = c("yes", "no"))

null_dist <- yrbss %>%
  specify(weight ~ physical_3plus) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "diff in means", order = c("yes", "no"))

ggplot(data = null_dist, aes(x = stat)) +
  geom_histogram()

```

### Exercise 6

How many of these null permutations have a difference of at least obs_stat?

-None of the null permutations have a difference of at least obs_stat.

```{r code-Exercise7}

null_dist %>%
  get_p_value(obs_stat = obs_diff, direction = "two_sided")

#sd
yrbss %>% 
  group_by(physical_3plus) %>% 
  summarise(sd_weight = sd(weight, na.rm = TRUE))

#mean
yrbss %>% 
  group_by(physical_3plus) %>% 
  summarise(mean_weight = mean(weight, na.rm = TRUE))

#sample size n
yrbss %>% 
  group_by(physical_3plus) %>% 
  summarise(freq = table(weight)) %>%
  summarise(n = sum(freq))

x0_3 <- 66.67389
n0_3 <- 4022
s0_3 <- 17.63805
x_3 <- 68.44847
n_3 <- 8342
s_3 <- 16.47832

z = 1.96

ub_0 <- x0_3 + z*(s0_3/sqrt(n0_3))
lb_0 <- x0_3 - z*(s0_3/sqrt(n0_3))
ub_0
lb_0

ub <- x_3 + z*(s_3/sqrt(n_3))
lb <- x_3 - z*(s_3/sqrt(n_3))

ub
lb
```

### Exercise 7

Construct and record a confidence interval for the difference between the weights of those who exercise at least three times a week and those who don’t, and interpret this interval in context of the data.

-The standard deviation is 17.63805 for students who do are not active at least 3 days per week and 16.47832 for those who are.

-The mean is 66.67389 for students who do are not active at least 3 days per week and 68.44847 for those who are.

-We are 95% confident that students who exercise at least three times a week have a mean weight between 68.09485 kg and 68.80209 kg. Students who don't exercise at least three times a week have a mean weight between 66.12878 kg and 67.219 kg with 95% confidence.


```{r code-Exercise8}

height_data <- yrbss %>% select(height) %>% na.omit()

meanheight <- mean(height_data$height)
sd3 <- sd(height_data$height)
max3 <- max(height_data$height)
sdheight <- sd(height_data$height)
stderrorheight <- sdheight / sqrt(nrow(height_data))

max3

meanheight + (2.5 * sd3)

tvalueheight <- qt(.05/2, nrow(height_data) - 1, lower.tail = FALSE)
rightintheight <- meanheight + tvalueheight * stderrorheight
leftintheight <- meanheight - tvalueheight * stderrorheight

leftintheight
rightintheight

```

### Exercise 8

Calculate a 95% confidence interval for the average height in meters (height) and interpret it in context.

-The max of height_data is 2.11 and the mean plus 2.5 sd is 1.952984. 
-The 95% confidence interval is from 1.689411 meters to 1.693071 meters.

```{r code-Exercise9}

tvalueheight <- qt(.1/2, nrow(height_data) - 1, lower.tail = FALSE)
rightintheight <- meanheight + tvalueheight * stderrorheight
leftintheight <- meanheight - tvalueheight * stderrorheight

leftintheight
rightintheight
```

### Exercise 9

Calculate a new confidence interval for the same parameter at the 90% confidence level. Comment on the width of this interval versus the one obtained in the previous exercise.

-The 90% confidence interval is from 1.69 meters to 1.693 meters 


```{r code-Exercise10}

height_exercise <- yrbss %>% 
  filter(physical_3plus == "yes") %>% 
  select(height) %>% 
  na.omit()

height_noexercise <- yrbss %>% 
  filter(physical_3plus == "no") %>% 
  select(height) %>% 
  na.omit()

# Box Plot
boxplot(height_exercise$height, height_noexercise$height,
        names = c("Exercise", "No Exercise"))


mean4 <- mean(height_noexercise$height)
sd4<- sd(height_noexercise$height)
max4 <- max(height_noexercise$height)

mean5 <- mean(height_exercise$height)
sd5 <- sd(height_exercise$height)
max5 <- max(height_exercise$height)

max4
mean4 + (2.5 * sd4)

max5
mean5 + (2.5 * sd5)

# Standard Error
meandiff <- mean5 - mean4
stderror <- 
  sqrt(
  ((mean5^2) / nrow(height_exercise)) +
  ((mean4^2) / nrow(height_noexercise))
  )

# T-Value and CI
degfreedomht2 <- 4022-1
tvalueht2 <- qt(.05/2, degfreedomht2, lower.tail = FALSE)
rightintervalht <- meandiff + tvalueht2 * stderror
leftintervalht <- meandiff - tvalueht2 * stderror

leftintervalht
rightintervalht

# P-Value
pvalueht2 <- 2*pt(tvalueht2,degfreedomht2, lower.tail = FALSE)

pvalueht2

```

### Exercise 10

Conduct a hypothesis test evaluating whether the average height is different for those who exercise at least three times a week and those who don’t.

-The max of height_noexercise is 2.11 and the mean plus 2.5 sd is 1.922732.

-The max of height_exercise is 2.11 and the mean plus 2.5 sd is 1.961452.

-The 95% confidence interval is from -0.02552409 to 0.1007759.

-The P-Value is 0.05...."If P is high, null will fly," so we fail to reject the null hypothesis.


```{r code-Exercise11}

yrbss %>% group_by(hours_tv_per_school_day) %>% summarise(n())

```

### Exercise 11

Now, a non-inference task: Determine the number of different options there are in the dataset for the hours_tv_per_school_day there are.

-There are 7 different options in the dataset hours_tv_per_school_day, not including those labeled "NA".


```{r code-Exercise12}

yrbss <- yrbss %>%
  mutate(sleep_less = ifelse(yrbss$school_night_hours_sleep < 6, "yes", "no"))
 
weight_less <- yrbss %>% 
  select(weight, sleep_less) %>% 
  filter(sleep_less == "yes") %>%
  na.omit()

weight_more <- yrbss %>% 
  select(weight, sleep_less) %>% 
  filter(sleep_less == "no") %>%
  na.omit()

boxplot(weight_less$weight, weight_more$weight,
        names = c("Less Sleep", "More Sleep"))

mn <- mean(weight_less$weight)
sd <- sd(weight_less$weight)
max <- max(weight_less$weight)
max

mn1 <- mean(weight_more$weight)
sd2 <- sd(weight_more$weight)
max2 <- max(weight_more$weight)

mean_diff <- mn1 - mn
sd <- 
  sqrt(
  ((mn1^2) / nrow(weight_more)) +
  ((mn^2) / nrow(weight_less))
  )

df <- 2492-1
t <- qt(.05/2, df, lower.tail = FALSE)

upper_ci <- mean_diff + t * sd
lower_ci <- mean_diff - t * sd

c(lower_ci ,upper_ci)


p_value <- 2*pt(t,df, lower.tail = FALSE)
p_value

```

### Exercise 12

Come up with a research question evaluating the relationship between height or weight and sleep. Formulate the question in a way that it can be answered using a hypothesis test and/or a confidence interval. Report the statistical results, and also provide an explanation in plain language. Be sure to check all assumptions, state your α level, and conclude in context.

-Question: Is there evidence that students who weigh more than the mean weight sleep more than students who are lighter than the mean weight?

Null: There is a relationship between weight and sleep

Alternative: There is no relationship between weight and sleep

95% confident level

Since the P-value equals alpha, we fail to reject the null hypothesis. We can't determine that a relationship exists between weight and sleep.